A Non-local Bayesian Video Denoising Method

### 2015-10-15 doc/201507_tip/paper.tex
[doc] Advances in TIP paper. Parameter section.

According to \cite{Vershynin2012}, the number of patches needed to achieve a given approximation
error between the empirical and the true covariances grows linearly with the
dimensionality of the space spanned by the data. 
\pcomment{Comment 1: Actually, I found results relating the number of samples with
the dimensionality of the ambient space. 
However, if the data lies in a $p$-dimensional subspace, one could 
compute a basis of this subspace from $n > p$ samples, and work in the low-dim
coordinates to estimate a covariance matrix there. In that way, one could build
a covariance matrix in the ambient space.}
\pcomment{Comment 2: This reference may not be as relevant, in the light of the 
results known for Gaussian covariance matrices. It would be interesting to see what 
\cite{bunea2015} says regarding covariance matrices with low-effective rank.}

The rank also plays a role in the estimation of the covariance matrix.
As explained in \S \ref{sec:complexity}, alleviating the computational cost is
one of the reasons for using a low rank approximation of the empirical
covariance matrix. But there are also good reasons regarding the denoising
performance. In situations in which the number of samples 
is comparable to their dimension, the eigen-decomposition of the empirical
covariance matrix $\widehat C$ can have considerable errors with respect to
that of the population covariance matrix $C$. 
%
The case of Gaussian models with covariance matrices of the form $C + \sigma^2I$,
where $C$ is of rank $r$, has been studied in the statistics and physics
literature. The authors in \cite{johnstone2009,paul2007} study the
limit when $n,p\to\infty$ such that $p/n\to\gamma < \infty$.
All these works show that, for a non-zero ratio $\gamma$, the empirical eigenvectors 
and eigenvalues are inconsistent estimators of the population ones.
Furthermore, the smaller eigenvalues and their eigenvectors will have a higher
error and eventually, if the signal-to-noise ratio drops bellow than $\sqrt{\gamma}$, 
the eigenvector is confounded with the noise and the resulting estimate will be,
in expectation, orthogonal to the actual value.
A similar behaviour was analysed by \cite{nadler2008} for finite $p,n$.
Therefore, when the dimension is high and comparable to the number of samples,
smaller eigenvalues and eigenvectors are very noisy.

@@

The plots in Figure \pcomment{TODO} show the PSNR of the basic estimate as a function of the 
rank $r_1$, for several values of $n_{1}$.
%
%
These results behave as expected. Smaller groups obtained by keeping only the
most similar patches have only a few modes of variation and thus are better
modelled with a low-rank covariance matrix. As the group of patches grows
bigger, patches in the group are not as similar anymore, and a higher rank is
needed to allow for additional modes of variation.

A similar qualitative behaviour can be observed for the second stage, as shown in Figure \pcomment{TODO}. 
The plots in the figure result from varying the 2nd stage $r_2$ and
$n_{2}$ parameters.
%
Note that for the 2nd stage best results are found with a much smaller value
for $n_{2}$ than in the first stage.
%
A possible explanation is that the noise-less patches, as well as the basic
estimates, lie close to a low-dimensional subspace. On the other hand, the
noisy patches, which have been perturbed by isotropic noise and span the entire
ambient space. Thus, following \cite{Vershynin2012}, more samples are required for
estimating the covariance matrix in the presence of noise.
\pcomment{Although we do not have an explanation about why higher levels of noise
require more samples.}

So far we have studied the parameters affecting the MAP estimation for each stage 
independently. However, to determine the optimal parameters, we need to look at
them jointly. The first stage parameters maximizing the PSNR of the basic
estimate, yield a suboptimal result in terms of the end PSNR.
%
Figure \pcomment{TODO} shows a correlation plot for the PSNR of the first and
second stages obtained when varying \pcomment{TODO}. It can be observed, that the 
results with the highest basic and final PSNR do not coincide.

The aims of the first and second stage are different: the output 
of the first stage is used as an oracle in the second stage. As such, it can
have more variance but it should have a small bias. The reason is that the
second stage is robust to some noise in the basic estimate, but
the signal components lost in the first stage are harder to recover.

Due the coupling of the first and second stages, determining the optimal values
for the pair ranks and number of patches requires a four dimensional parameter
search, which would take a lot of time.
%
Instead, we fix the first stage parameters at
\[r_1 = s_{x,1}^2s_{t,1}\quad\quad n_{1} = 400,\]
which we found to give good results. We are being conservative and use the
full rank covariance matrix to eliminate the least signal components.

We then run a parameter search to determine $r_2$ and $n_{2}$.
The result is the one in Figure \pcomment{figure used before}. From here
establish the following values as a function of $\sigma$:
\[r_2(\sigma) = 40 - \frac13(\sigma - 10)\quad\quad n_{2} = 150.\]
Note that smaller value for $n_{2}$ could give better results,
but this comes at a higher computational cost.

@@

\begin{figure*}[htpb!]
	\begin{center}
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s10_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s20_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s30_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s40_average_derf.eps}\\
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-image_s10_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-image_s20_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-image_s30_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-image_s40_average_mono.eps}\\
	\end{center}
	\caption{Top row: average PSNRs over four sequences, obtained by varying the
		second stage parameters $r$ and $n$. From left to right, results for
		$\sigma = 10, 20, 40$. In each image, the contrast has been stretched so
		that white corresponds to the maximum PSNR and black to a decay of 1dB
		bellow the maximum. 
		Bottom: same for grayscale sequences.} 
	\label{fig:rank-nsim-image-stage2}
\end{figure*}

\begin{figure*}[htpb!]
	\begin{center}
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s10_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s20_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s30_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s40_average_derf.eps}\\
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s10_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s20_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s30_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-image_s40_average_mono.eps}\\
	\end{center}
	\caption{Same as Figure \ref{fig:rank-nsim-image-stage2} but for the first stage.}
	\label{fig:rank-nsim-image-stage1}
\end{figure*}

>>>>

The problem of estimating the covariance matrix has received a lot of attention in the
past ten years.
According to \cite{Vershynin2012}, the number of samples needed to achieve a given approximation
error between the empirical and the true covariances grows linearly with the
dimensionality of the space spanned by the data. 
%\pcomment{Comment: Actually, I found results relating the number of samples with
%the dimensionality of the ambient space. 
%However, if the data lies in a $p$-dimensional subspace, one could 
%compute a basis of this subspace from $n > p$ samples, and work in the low-dim
%coordinates to estimate a covariance matrix there. In that way, one could build
%a covariance matrix in the ambient space.}
\pcomment{Comment: This reference may not be as relevant, in the light of the 
results known for Gaussian covariance matrices. It would be interesting to see what 
\cite{bunea2015} says regarding covariance matrices with low-effective rank.}

The rank also plays a role in the estimation of the covariance matrix.
As explained in \S \ref{sec:complexity}, alleviating the computational cost is
one of the reasons for using a low rank approximation of the empirical
covariance matrix. But there are also good reasons regarding the denoising
performance. In situations in which the number of samples 
is comparable to their dimension, the eigen-decomposition of the empirical
covariance matrix $\widehat C$ can have considerable errors with respect to
that of the population covariance matrix $C$. 
%
The case of Gaussian models with covariance matrices of the form $C = A + \sigma^2I$,
where $A$ is does not have full rank, has been studied in the statistics and physics
literature. The authors in \cite{johnstone2009,paul2007,nadler2008} study the
limit when $n,p\to\infty$ such that $p/n\to\gamma < \infty$.
These works show that, for a non-zero ratio $\gamma$, the empirical eigenvectors 
and eigenvalues are inconsistent estimators of the population ones.
Furthermore, the smaller eigenvalues and their eigenvectors will have a higher
error and eventually, if the signal-to-noise ratio drops bellow $\sqrt{\gamma}$, 
the eigenvector is confused with the noise and the resulting estimate will be,
in expectation, orthogonal to the actual value.
Finite sample bounds for the eigenvalues and eigenvectors have been also studied 
by \cite{nadler2008}, in the case in which $A$ has rank 1.
\pcomment{What can we say about this bounds.}
These larger errors between the sample eigenvectors and their population
counterparts manifest themselves as noise, which then affects the MAP estimate.
This suggest that better denoising should be obtained by discarding the
eigenvectors associated with lower eigenvalues.

@@

The plots in Figure \ref{fig:rank1-nsim1-bpsnr} show the PSNR of the basic
estimate as a function of the rank $r_1$, for several values of $n_{1}$.
The curves correspond to an average over 20 frames clips 
extracted from four classical test
sequences, namely \emph{bus}, \emph{foreman}, \emph{football} and \emph{tennis}.
The sequences were contaminated with AGWN of standard deviation $\sigma = 10,
20, 30, 40$. 
%
%
We can interpret the results in the light of the theoretical results
described above.
For all noise levels, best results are obtained when using a larger number of
samples $n_1$. All curves of fixed $n_1$ show a maximum at a given rank lower than
the dimensionality of the patch. As the noise level increases, this optimal rank
shifts to the left. This makes sense, since less eigenvectors are reliable estimates.
%
Also, for a fixed noise level, the optimal rank is smaller for smaller values of $n_1$.
Smaller groups obtained by keeping only the
most similar patches have less modes of variation, and thus a worse signal-to-noise ratio
in the eigenvectors with lower eigenvalues.
%As the group of patches grows
%bigger, patches in the group are not as similar anymore, and a higher rank is
%needed to allow for additional modes of variation.

The results obtained for the final estimate when varying the second stage
parameters $n_2$ and $r_2$ show a different behaviour, as shown in Figure
\ref{fig:rank2-nsim2-fpsnr}.
These curves where obtained fixing the first stage parameters at $n_1 = 400$
and $r_1 = 100$ (full rank).
\pcomment{It doesn't make much sense considering $r_1 = 100$ after saying that
is best to consider a low $r_1$.}
%
In the second stage best results are found with a much smaller value
for $n_{2}$ than in the first stage.
%
The reason for this is that in the second stage the sample covariance is computed
using patches from the basic estimate, with a much lower noise level.
In this setting, there is no gain in increasing the number of patches,
and there seems to be a loss possibly due to the fact that smaller groups of patches
are better modelled by a Gaussian distribution.
\pcomment{I don't have an explanation for the curves for $\sigma = 10$. Precisely
in the lower noise level, the PSNR drops when $n_2$ is reduced, except for lower ranks.
This contradicts the argument given: for less noise, it should be always better to consider
less patches, not only for lower ranks.}

So far we have studied the parameters affecting the MAP estimation for each stage 
independently. However, to determine the optimal parameters, we need to look at
them jointly. The first stage parameters maximizing the PSNR of the basic
estimate, yield a suboptimal result in terms of the end PSNR.
%
Figure \ref{fig:rank1-nsim1-bpsnr-fpsnr} plots for the PSNR of the first and
second stages obtained when varying $n_1$ and $r_1$ as in Figure
\ref{fig:rank1-nsim1-bpsnr}. It can be observed, that the 
results with the highest basic and final PSNR do not coincide.
%
The aims of the first and second stage are different: the output 
of the first stage is used as an oracle in the second stage. As such, it can
have more variance but it should have a small bias. The reason is that the
second stage is robust to some noise in the basic estimate, but
the signal components lost in the first stage are harder to recover.
\pcomment{The figure shows actually the opposite!!}

Due the coupling of the first and second stages, determining the optimal values
for the ranks $r_1, r_2$ and the number of patches $n_1,n_2$ requires a four
dimensional parameter search.
%
Instead, we fix the first stage parameters at
\[r_1 = s_{x,1}^2s_{t,1}\quad\quad n_{1} = 400,\]
which we found to give good results. In doing so, we are being conservative and
use the full rank covariance matrix to eliminate the least signal components.
%
We then run a parameter search to determine $r_2$ and $n_{2}$.
The results is the ones shown in Figure \ref{fig:rank1-nsim1-fpsnr}. From here
establish the following values as a function of $\sigma$:
%\[r_2(\sigma) = 40 - \frac13(\sigma - 10)\quad\quad n_{2} = 150.\]
\[r_2(\sigma) = 35 - \frac12(\sigma - 10)\quad\quad n_{2} = 100.\]
Note that a higher value for $n_{2}$ would give slightly worse results,
at a lower computational cost.

@@

\begin{figure*}[htpb!]
	\begin{center}
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s10_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s20_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s30_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s40_average_derf.eps}\\

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

	\end{center}
	\caption{Basic estimate PSNR, averaged over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. In the top row: results for RGB sequences, bottom
		row, results for grayscale sequences.} 
	\label{fig:rank1-nsim1-bpsnr}
\end{figure*}

\begin{figure*}[htpb!]
	\begin{center}
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s10_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s20_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s30_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s40_average_derf.eps}\\

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

	\end{center}
	\caption{PSNRs of final vs. basic estimate. Averages over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. Top: results for RGB sequences; bottom:
		results for grayscale sequences. The width of the line is proportional to
		the rank. \pcomment{These plots contradict the argument in favour of using
		full rank in the first stage (supposedly introducing a higher variance, but a
		smaller bias). It turns out, that (1) the final PSNR is pretty much
		correlated with the basic PSNR, and (2) the optimal rank for the final
		estimate is lower than that of the basic estimate. I do not have an explanation for this.}} 
	\label{fig:rank1-nsim1-bpsnr-fpsnr}
\end{figure*}

\begin{figure*}[htpb!]
	\begin{center}
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s10_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s20_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s30_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s40_average_derf.eps}\\

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

	\end{center}
	\caption{Final estimate PSNR, averaged over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. In the top row: results for RGB sequences, bottom
		row, results for grayscale sequences.} 
	\label{fig:rank1-nsim1-fpsnr}
\end{figure*}

\begin{figure*}[htpb!]
	\begin{center}
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s10_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s20_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s30_average_derf.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s40_average_derf.eps}\\

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

	\end{center}
	\caption{Final estimate PSNR, averaged over four sequences, when varying the
		second stage parameters $r_2$ and $n_2$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. In the top row: results for RGB sequences, bottom
		row, results for grayscale sequences.} 
	\label{fig:rank2-nsim2-fpsnr}
\end{figure*}
